---
title: "Evolution Canyon"
author: "Jay T. Lennon, Kenneth J. Locey, & Mario E. Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output:
  pdf_document:
    fig_caption: true
---

Dormancy, dispersal, and environmental filtering at "Evolution Canyon"

Brief Description


# Initial Setup
```{r results='hide', message=FALSE}
rm(list=ls())
getwd()
setwd("~/GitHub/evolution-canyon/analyses")

# Import Required Packages
require("png")
require("grid")
require("vegan")
```


# Figure 1: Conceptual model/theory and predictions of how environmental filtering and dormancy influence community assembly (with dispersal being implicit and more or less high) 

**Jay is updating this**

```{r fig.width=2, fig.height=4,echo=FALSE,fig.cap="Figure 1: Conceptual Model"}
img <- readPNG("../figures/ConceptualModel.png")
grid.raster(img)
```


# Figure 2: Photo of EC and simulation of EC-like landscape

```{r fig.width=2, fig.height=4,echo=FALSE,fig.cap="Figure 1: Conceptual Model"}
img1 <- readPNG("../figures/ECLandscape.png")
img2 <- readPNG("../figures/BIDELandscape.png")
grid.newpage()
grid.raster(img1, x=0.5, y = 0.75, width=0.5)
grid.raster(img2, x = 0.5, y = 0.35, width=0.5)
```

# Figure 3: BIDE Model Ordinations

Ordination of simulated data capturing processes/scenarios described in Fig. 1. Here, we only have models 1, 2, 3, and 6; missing models 4 and 5. Can these other scenarios be captures with simplex?


```{r}
source("../bin/ec.pcoa.R")
input <- list.files("../microbide/SbyS/", "*.txt")

# Create Blank Output
pcoa.output <- list()

# Run PCoA On All Samples
for (i in 1:length(input)){
  pcoa.output[[input[i]]] = ec.pcoa.fun(shared = paste("../microbide/SbyS/", 
                                             input[i], sep=""), 
                                             cutoff = "0.03", 
                                             design = "../data/simmy.design.txt")
  }

layout(matrix(1:6, 3, 2, byrow=T), width = 3, height = 3)

# Plot Each
for (i in 1:length(pcoa.output)){
  
  
  # Plot Parameters
  par(mfrow=c(1,1), mar=c(5,5,1,1))
  layout(rbind(1, 2), height=c(7, 1))
  x.dim <- c(min(pcoap$V2)+min(pcoap$V2)*0.2,max(pcoap$V2)+max(pcoap$V2)*0.2)
  y.dim <- c(min(pcoap$V1)+min(pcoap$V1)*0.2,max(pcoap$V1)+max(pcoap$V1)*0.2)

  # Initiate Plot
  plot(pcoap$V2, pcoap$V1, xlab=paste("PCoA Axis 2 (",explainvar2, "%)", sep="")
    , ylab=paste("PCoA Axis 1 (",explainvar1, "%)", sep=""),
    xlim=x.dim,ylim= y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)
  axis(side=1, las=1)
  axis(side=2, las=1)
  abline(h=0, lty="dotted")
  abline(v=0, lty="dotted")
  mol.shape <- rep(NA, dim(pcoap)[1])
    for (i in 1:length(mol.shape)){
      if (pcoap$molecule[i] == "DNA"){mol.shape[i] = 21}
      else {mol.shape[i] = 22}
      }
  slope.color <- rep(NA, dim(pcoap)[1])
    for (i in 1:length(slope.color)){
      if (pcoap$slope[i] == levels(pcoap$slope)[1]) {slope.color[i] = "brown"}
      else {slope.color[i] = "green3"}
      }
  points(pcoap$V2, pcoap$V1, pch=mol.shape, cex=2.0, col="black", bg=slope.color, lwd=2)
  ordiellipse(cbind(pcoap$V2, pcoap$V1), pcoap$slope.molecule, kind="sd", conf=0.95,
    lwd=2, lty=3, draw = "lines", col = "black", label=TRUE)
  # legend("topleft", c(paste("All; ",levels(pcoap$slope)[1]," Slope", sep=""),
  #   paste("All; ",levels(pcoap$slope)[2]," Slope", sep=""),
  #   paste("Active; ",levels(pcoap$slope)[1]," Slope", sep=""),
  #   paste("Active; ",levels(pcoap$slope)[2]," Slope", sep="")),
  #   pt.lwd=2, col="black", pt.bg=c("brown3", "green3", "brown3",
  #   "green3"), pch=c(21,21,22,22), bty='o', box.lty=0, bg="white", cex=1.5)
  box(lwd=2)
  par(mar=c(0, 3, 0, 0))
  plot.new()
  legend("center", c(paste("All; ",levels(pcoap$slope)[1]," Slope", sep=""),
    paste("All; ",levels(pcoap$slope)[2]," Slope", sep=""),
    paste("Active; ",levels(pcoap$slope)[1]," Slope", sep=""),
    paste("Active; ",levels(pcoap$slope)[2]," Slope", sep="")),
    pt.lwd=2, col="black", pt.bg=c("brown", "green3", "brown",
    "green3"), pch=c(21,21,22,22), bty='n', ncol=2, cex=1.5, pt.cex=2)




  
  
  
}

```

```{r}


```

# Multivariate Analysis of BIDE Simulations

A general discussion that the simulations agreed with theory outlined in Fig. 1. 
Will reference ordinations and PCoA. 
Discussed Table 1 having summary data from the PERMANOVA results.

```{r}
input <- list.files("../microbide/SbyS/", "*.txt")

permanova.output <- list()

require(vegan)

for (i in c(1, 3:length(input))){
  shared <- paste("../microbide/SbyS/", input[i], sep="")
  design <- "../data/simmy.design.txt"
  sbys.in <- read.otu(shared, "0.03")
  design.in <- read.delim(design, header=T, row.names=1)
  
  # Remove any OTUs with <2 OTUs
  sbys.red <- ec_data[which(colSums(sbys.in) >= 2),]
  
  # Relative Abundances
  sbysREL <- sbys.red
  for(j in 1:nrow(sbys.red)){
    sbysREL[j,] = sbys.red[j,]/sum(sbys.red[j,])
  }
  
  
  # Log Transformation
  sbysREL.log <- decostand(sbysREL, method="log")

  Adonis.b <- adonis(sbysREL ~ design.in$molecule * design.in$slope)
  
  permanova.output[[input[i]]] = Adonis.b
  
}



```

# Figure 4: Emperical Data

Ordination of empirical data consistent with the view that non-random entrance and exiting from dormancy on top of environmental filtering influence community assembly

```{r}



```

