---
title: "Evolution Canyon"
author: "Mario E. Muscarella, Kenneth J. Locey, & Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output:
  pdf_document:
  fig_caption: true
---

Dormancy, dispersal, and environmental filtering at "Evolution Canyon"

Brief Description:


# Initial Setup
```{r results='hide', message=FALSE}
rm(list=ls())
getwd()
setwd("~/GitHub/evolution-canyon/analyses")

# Import Required Packages
require("png")
require("grid")
require("vegan")


# Import Scripts
source("../bin/DiversityFunctions.r")
source("../bin/ec.dbRDA.R")

# Common Functions
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```

# BIDE Model Parameter Exploration
```{r}
dir <- "../microbide/SbyS/supp"
input <- list.dirs(dir, full.names = T)

# Remove runs without all conditions
missing <- 1
for(i in 2:length(input)){
  if("Condition1.txt" %in% list.files(input[i]) == FALSE | 
     "Condition4.txt" %in% list.files(input[i]) == FALSE){
    missing <- c(missing, i)
  }
}
input <- input[-missing]

names <- unlist(lapply(strsplit(input, "supp/"), "[[", 2))
names.long <- data.frame(do.call('rbind', strsplit(names, "_", fixed = T)))

lgp <- as.numeric(gsub("lgp=", "", names.long[, 1]))
im <- as.numeric(gsub("im=", "", names.long[, 2]))
dk <- as.numeric(gsub("dispkernel=", "", names.long[, 3]))

# Conditions
cond1 <- paste(input, "Condition1.txt", sep = "/")
cond4 <- paste(input, "Condition4.txt", sep = "/")
```

# Pair Distances
```{r}
# Define Function
pair.dis <- function(x, rarefy = FALSE){
  sbys <- read.otu(x, "0.03")
  if (rarefy){
    sbys <- rrarefy(sbys, min(rowSums(sbys)))
  } else {}
  grep("north", rownames(sbys))
  grep("south", rownames(sbys))
  grep("all", rownames(sbys))
  grep("active", rownames(sbys))
  slope <- gsub("[0-9]*_.*", "", rownames(sbys))
  molecule <- gsub(".*_", "", rownames(sbys))
  pair <- gsub("_.*", "", gsub(".*th", "", rownames(sbys)))
  
  # Remove OTUs with < 1 global occurence
  if(min(colSums(sbys)) < 2){
    sbys <- sbys[, -c(which(colSums(sbys) < 2))]
  }
  
  # Calculating Relative Abundance
  dataREL <- sbys
  for(i in 1:nrow(sbys)){
    dataREL[i,] = sbys[i,]/sum(sbys[i,])
  }
  
  # Log Transform Relative Abundance
  dataREL.l <- suppressWarnings(decostand(dataREL,method="log"))
  
  # Distance Based Redundancy Analysis
  dbRDA <- capscale(dataREL.l ~ slope + molecule , distance="bray", warn = F)
  
  # Calculate BC Dist
  bide.dist <- vegdist(dataREL.l, method = "bray", binary = "FALSE")
  
  # Principal Coordinates Analysis
  bide.pcoa <- cmdscale(bide.dist, eig = TRUE, k = 5)
  
  # Distance Between Pairs
  pair.m <- as.matrix(bide.dist)[seq(1, 80, by = 2), seq(2, 80, by = 2)]
  pair.dist <- diag(pair.m)
  pair.dist.slope <- gsub("[0-9]*_.*", "", row.names(pair.m))
  m <- aggregate(pair.dist, by = list(pair.dist.slope), mean)
  s <- aggregate(pair.dist, by = list(pair.dist.slope), sem)
  out <- data.frame("north.mean" = m[1, 2], "north.sem" = s[1, 2],
         "south.mean" = m[2, 2], "south.sem" = s[2,2])
  return(out)
}

# Condition 1
pair.distances1 <- as.data.frame(matrix(NA, nrow = length(cond1), ncol = 7))
colnames(pair.distances1) <- c("lgp", "im", "dk", "north.mean", "north.sem", 
                              "south.mean", "south.sem")

pair.distances1$lgp <- lgp; pair.distances1$im <- im; pair.distances1$dk <- dk

for(i in 1:length(cond1)){
  pair.distances1[i, 4:7] <- pair.dis(cond1[i], rarefy = TRUE)
}

# Condition 4
pair.distances4 <- as.data.frame(matrix(NA, nrow = length(cond4), ncol = 7))
colnames(pair.distances4) <- c("lgp", "im", "dk", "north.mean", "north.sem", 
                              "south.mean", "south.sem")
pair.distances4$lgp <- lgp; pair.distances4$im <- im; pair.distances4$dk <- dk

for(i in 1:length(cond4)){
  pair.distances4[i, 4:7] <- pair.dis(cond4[i], rarefy = TRUE)
}
```

# Plot
```{r}
# Initialize Plot

png(filename = "../figures/BIDE_PairDistance.png", 
    width = 1800, height = 1200, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0.5,0.5), oma=c(5, 8, 1, 1))
layout(matrix(c(1:6), 2, 3, byrow=T))

# Reorganize Condition 1 for Plotting
pair.dis.b <- data.frame(
  lgp = rep(pair.distances1$lgp, 2),
  im = rep(pair.distances1$im, 2),
  dk = rep(pair.distances1$dk, 2),
  dist = c(pair.distances1$north.mean, pair.distances1$south.mean), 
  col = as.character(c(rep("sienna", length(pair.distances1$north.mean)), 
          rep("forestgreen", length(pair.distances1$south.mean)))),
  stringsAsFactors = F)
pair.dis.b <- pair.dis.b[sample(1:dim(pair.dis.b)[1]),]

# Plot Condition 1
plot(pair.dis.b$dist ~ pair.dis.b$lgp, col = pair.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = F, las = 1, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = T, las = 1, cex.axis = 1.25)
mtext("Pair Dissimilarity", side = 2, line = 3.5, outer = F)
mtext("Condition 1", side = 2, line = 5.5, outer = F)
box(lwd = 1.5)

x1 <- pair.distances1$lgp; x1.o <- order(x1)
loess.50 <- loess(pair.distances1$north.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(pair.distances1$south.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)

legend("topleft", legend = c("North", "South"),  
       col = c("sienna", "forestgreen"),
       lty = 1, bty = "n")

plot(pair.dis.b$dist ~ pair.dis.b$im, col = pair.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = F, las = 1, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = F, las = 1)
box(lwd = 1.5)


x1 <- pair.distances1$im; x1.o <- order(x1)
loess.50 <- loess(pair.distances1$north.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(pair.distances1$south.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)


plot(pair.dis.b$dist ~ log10(pair.dis.b$dk), col = pair.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = F, las = 1, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = F, las = 1)
box(lwd = 1.5)

x1 <- log10(pair.distances1$dk); x1.o <- order(x1)
loess.50 <- loess(pair.distances1$north.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(pair.distances1$south.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)


# Reorganize Condition 4 for Plotting
pair.dis.b <- data.frame(
  lgp = rep(pair.distances4$lgp, 2),
  im = rep(pair.distances4$im, 2),
  dk = rep(pair.distances4$dk, 2),
  dist = c(pair.distances4$north.mean, pair.distances4$south.mean), 
  col = as.character(c(rep("sienna", length(pair.distances4$north.mean)), 
          rep("forestgreen", length(pair.distances4$south.mean)))),
  stringsAsFactors = F)
pair.dis.b <- pair.dis.b[sample(1:dim(pair.dis.b)[1]),]

# Plot Condition 4
plot(pair.dis.b$dist ~ pair.dis.b$lgp, col = pair.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = T, las = 1, cex.axis = 1.25, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = T, las = 1, cex.axis = 1.25)
mtext("Pair Dissimilarity", side = 2, line = 3.5, outer = F)
mtext("Condition 4", side = 2, line = 5.5, outer = F)
mtext("Logseries Parameter", side = 1, line = 3, outer = F)
box(lwd = 1.5)

x1 <- pair.distances4$lgp; x1.o <- order(x1)
loess.50 <- loess(pair.distances4$north.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(pair.distances4$south.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)

plot(pair.dis.b$dist ~ pair.dis.b$im, col = pair.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = T, las = 1, cex.axis = 1.25, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = F, las = 1)
mtext("Immigration Rate", side = 1, line = 3, outer = F)
box(lwd = 1.5)

x1 <- pair.distances4$im; x1.o <- order(x1)
loess.50 <- loess(pair.distances4$north.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(pair.distances4$south.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)



plot(pair.dis.b$dist ~ log10(pair.dis.b$dk), col = pair.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = T, las = 1, cex.axis = 1.25, at = c(-3, -2, -1))
axis(side = 2, labels = F, las = 1)
mtext(expression(paste("Log"[10], " (Dispersal Density Kernel)")), 
      side = 1, line = 3, outer = F)
box(lwd = 1.5)

x1 <- log10(pair.distances4$dk); x1.o <- order(x1)
loess.50 <- loess(pair.distances4$north.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(pair.distances4$south.mean[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)


# Close Plot Device
dev.off()
graphics.off()
#```

## Show Plot
#```{r}
img <- readPNG("../figures/BIDE_PairDistance.png")
grid.raster(img)

```

# Site Distances
```{r}
site.dis <- function(x, rarefy = FALSE){
  sbys <- read.otu(x, "0.03")
  if (rarefy){
    sbys <- rrarefy(sbys, min(rowSums(sbys)))
  } else {}
  grep("north", rownames(sbys))
  grep("south", rownames(sbys))
  grep("all", rownames(sbys))
  grep("active", rownames(sbys))
  slope <- gsub("[0-9]*_.*", "", rownames(sbys))
  molecule <- gsub(".*_", "", rownames(sbys))
  pair <- gsub("_.*", "", gsub(".*th", "", rownames(sbys)))
  
  # Remove OTUs with < 1 global occurence
  if(min(colSums(sbys)) < 2){
    sbys <- sbys[, -c(which(colSums(sbys) < 2))]
  }
  # Calculating Relative Abundance
  dataREL <- sbys
  for(i in 1:nrow(sbys)){
    dataREL[i,] = sbys[i,]/sum(sbys[i,])
  }
  
  # Log Transform Relative Abundance
  dataREL.l <- suppressWarnings(decostand(dataREL,method="log"))
  
  # Distance Based Redundancy Analysis
  dbRDA <- capscale(dataREL.l ~ slope + molecule , distance="bray", warn = F)
  
  # Calculate BC Dist
  bide.dist <- vegdist(dataREL.l, method = "bray", binary = "FALSE")
  
  # Principal Coordinates Analysis
  bide.pcoa <- cmdscale(bide.dist, eig = TRUE, k = 5)
  
  # Beta Dispersion
  grouping <- paste(slope, molecule, sep = "_")
  bide.disper <- betadisper(d = bide.dist, group = grouping)
  
  full.dis <- dist(bide.disper$centroids[, 1:5], method = "euclid", diag = T)
  
  # Distance Between Sites_All
  site.dist_all <- as.matrix(full.dis)["north_all", "south_all"]
  site.dist_active <- as.matrix(full.dis)["north_active", "south_active"]
  site.dist_north <- as.matrix(full.dis)["north_all", "north_active"]
  site.dist_south <- as.matrix(full.dis)["south_all", "south_active"]

  out <- data.frame("north" = site.dist_north, "south" = site.dist_south,
         "all" = site.dist_all, "active" = site.dist_active)
  return(out)
}


# Condition 1
site.distances1 <- as.data.frame(matrix(NA, nrow = length(cond1), ncol = 7))
colnames(site.distances1) <- c("lgp", "im", "dk", "north", "south", 
                              "all", "active")
site.distances1$lgp <- lgp; site.distances1$im <- im; site.distances1$dk <- dk

for(i in 1:length(cond1)){
  site.distances1[i, 4:7] <- site.dis(cond1[i], rarefy = T)
}

# Condition 4
site.distances4 <- as.data.frame(matrix(NA, nrow = length(cond4), ncol = 7))
colnames(site.distances4) <- c("lgp", "im", "dk", "north", "south", 
                              "all", "active")
site.distances4$lgp <- lgp; site.distances4$im <- im; site.distances4$dk <- dk

for(i in 1:length(cond4)){
  site.distances4[i, 4:7] <- site.dis(cond4[i],  rarefy = T)
}
```

```{r}

png(filename = "../figures/BIDE_SiteDistance.png", 
    width = 1800, height = 1200, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0.5,0.5), oma=c(5, 8, 1, 1))
layout(matrix(c(1:6), 2, 3, byrow=T))

# Reorganize Condition 1 for Plotting
site.dis.b <- data.frame(
  lgp = rep(site.distances1$lgp, 4),
  im = rep(site.distances1$im, 4),
  dk = rep(site.distances1$dk, 4),
  dist = c(site.distances1$north, site.distances1$south,
           site.distances1$active, site.distances1$all), 
  col = as.character(c(rep("sienna", length(site.distances1$north)), 
          rep("forestgreen", length(site.distances1$south)),
          rep("cornflowerblue", length(site.distances1$active)),
          rep("magenta", length(site.distances1$all)))),
  stringsAsFactors = F)
site.dis.b <- site.dis.b[sample(1:dim(site.dis.b)[1]),]


plot(site.dis.b$dist ~ site.dis.b$lgp, col = site.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = F, las = 1, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = T, las = 1, cex.axis = 1.25)
mtext("Site Dissimilarity", side = 2, line = 3.5, outer = F)
mtext("Condition 1", side = 2, line = 5.5, outer = F)
box(lwd = 1.5)

x1 <- site.distances1$lgp; x1.o <- order(x1)
loess.50 <- loess(site.distances1$north[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(site.distances1$south[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)
loess.50 <- loess(site.distances1$active[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="cornflowerblue", lwd = 2)
loess.50 <- loess(site.distances1$all[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="magenta", lwd = 2)

legend("topleft", legend = c("North", "South", "Active", "All"),  
       col = c("sienna", "forestgreen", "cornflowerblue", "magenta"),
       lty = 1, bty = "n")

plot(site.dis.b$dist ~ site.dis.b$im, col = site.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = F, las = 1, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = F, las = 1)
box(lwd = 1.5)

x1 <- site.distances1$im; x1.o <- order(x1)
loess.50 <- loess(site.distances1$north[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(site.distances1$south[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)
loess.50 <- loess(site.distances1$active[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="cornflowerblue", lwd = 2)
loess.50 <- loess(site.distances1$all[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="magenta", lwd = 2)

plot(site.dis.b$dist ~ log10(site.dis.b$dk), col = site.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = F, las = 1, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = F, las = 1)
box(lwd = 1.5)

x1 <- log10(site.distances1$dk); x1.o <- order(x1)
loess.50 <- loess(site.distances1$north[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(site.distances1$south[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)
loess.50 <- loess(site.distances1$active[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="cornflowerblue", lwd = 2)
loess.50 <- loess(site.distances1$all[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="magenta", lwd = 2)

# Reorganize Condition 4 for Plotting
site.dis.b <- data.frame(
  lgp = rep(site.distances4$lgp, 4),
  im = rep(site.distances4$im, 4),
  dk = rep(site.distances4$dk, 4),
  dist = c(site.distances4$north, site.distances4$south,
           site.distances4$active, site.distances4$all), 
  col = as.character(c(rep("sienna", length(site.distances4$north)), 
          rep("forestgreen", length(site.distances4$south)),
          rep("cornflowerblue", length(site.distances4$active)),
          rep("magenta", length(site.distances4$all)))),
  stringsAsFactors = F)
site.dis.b <- site.dis.b[sample(1:dim(site.dis.b)[1]),]

plot(site.dis.b$dist ~ site.dis.b$lgp, col = site.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = T, las = 1, cex.axis = 1.25, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = T, las = 1, cex.axis = 1.25)
mtext("Site Dissimilarity", side = 2, line = 3.5, outer = F)
mtext("Condition 4", side = 2, line = 5.5, outer = F)
mtext("Logseries Parameter", side = 1, line = 3, outer = F)
box(lwd = 1.5)

x1 <- site.distances4$lgp; x1.o <- order(x1)
loess.50 <- loess(site.distances4$north[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(site.distances4$south[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)
loess.50 <- loess(site.distances4$active[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="cornflowerblue", lwd = 2)
loess.50 <- loess(site.distances4$all[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="magenta", lwd = 2)

plot(site.dis.b$dist ~ site.dis.b$im, col = site.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = T, las = 1, cex.axis = 1.25, at = c(0.8, 0.9, 1.0))
axis(side = 2, labels = F, las = 1)
mtext("Immigration Rate", side = 1, line = 3, outer = F)
box(lwd = 1.5)

x1 <- site.distances4$im; x1.o <- order(x1)
loess.50 <- loess(site.distances4$north[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(site.distances4$south[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)
loess.50 <- loess(site.distances4$active[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="cornflowerblue", lwd = 2)
loess.50 <- loess(site.distances4$all[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="magenta", lwd = 2)

plot(site.dis.b$dist ~ log10(site.dis.b$dk), col = site.dis.b$col, pch = 16, 
     ylim = c(0, 0.2), las = 1, cex = 0.75, xlab = "", ylab = "", axes = F)
axis(side = 1, labels = T, las = 1, cex.axis = 1.25, at = c(-3, -2, -1))
axis(side = 2, labels = F, las = 1)
mtext(expression(paste("Log"[10], " (Dispersal Density Kernel)")), 
      side = 1, line = 3, outer = F)
box(lwd = 1.5)

x1 <- log10(site.distances4$dk); x1.o <- order(x1)
loess.50 <- loess(site.distances4$north[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="sienna", lwd = 2)
loess.50 <- loess(site.distances4$south[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="forestgreen", lwd = 2)
loess.50 <- loess(site.distances4$active[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="cornflowerblue", lwd = 2)
loess.50 <- loess(site.distances4$all[x1.o] ~ x1[x1.o], span=0.5) 
smoothed.50 <- predict(loess.50) 
lines(smoothed.50, x=sort(x1), col="magenta", lwd = 2)



# Close Plot Device
dev.off()
graphics.off()
#```

## Show Plot
#```{r}
img <- readPNG("../figures/BIDE_SiteDistance.png")
grid.raster(img)

```








input
test <- read.delim(paste(dir, input[1], sep = ""), header = T)
rowSums(as.matrix(test[ 1, -c(1:3)]))

# Create Blank Output
dbRDA.output <- list()

# Run PCoA On All Samples
for (i in 1:length(input)){
  dbRDA.output[[input[i]]] = ec.dbRDA.sim(shared = paste(dir, input[i], sep=""), 
                                             level = "0.03", 
                                             design = "../data/simmy.design.txt")
}

png(filename = "../figures/BIDE_Ordination.png", 
    width = 1600, height = 1200, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0,0), oma=c(6, 5, 1, 1))
layout(matrix(c(5, 5, 1:4), 3, 2, byrow=T), heights=c(1, rep(5, 2)))

labs <- c("A", "B", "C", "D", "E", "F")
x.lab <- c(F, F, T, T)
y.lab <- c(T, F, T, F)

# Plot Each
for (p in 1:length(dbRDA.output)){
  RDA <- dbRDA.output[[p]]$RDA
  explainvar1 <- dbRDA.output[[p]]$explainvar1
  explainvar2 <- dbRDA.output[[p]]$explainvar2
  print(explainvar1)
  print(explainvar2)

  x.dim <- c(-0.6, 0.6)
  y.dim <- c(-0.5, 0.5)

  # Initiate Plot
  plot(RDA$CAP1, RDA$CAP2, xlab = "", ylab="",
    xlim=x.dim,ylim= y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)

  axis(side=1, las=1, label=x.lab[p], lwd.ticks = 2, cex.axis = 1.25)
  axis(side=2, las=1, label=y.lab[p], lwd.ticks = 2, cex.axis = 1.25)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  abline(h=0, lty="dotted")
  abline(v=0, lty="dotted")
  mol.shape <- rep(NA, dim(RDA)[1])
    for (i in 1:length(mol.shape)){
      if (RDA$molecule[i] == "DNA"){mol.shape[i] = 21}
      else {mol.shape[i] = 22}
      }
  slope.color <- rep(NA, dim(RDA)[1])
    for (i in 1:length(slope.color)){
      if (RDA$slope[i] == levels(RDA$slope)[1]) {slope.color[i] = "wheat3"}
      else {slope.color[i] = "green3"}
      }
  points(RDA$CAP1, RDA$CAP2, pch=mol.shape, cex=2.0, col="black", bg=slope.color, lwd=2)
  box(lwd=2)
  
  text(x = -0.58, y = 0.45, labs[p], cex = 1.5)
}
  
mtext("Axis 1", side = 1, cex = 1.5, outer=T, line = 3)
mtext("Axis 2", side = 2, cex = 1.5, outer=T, line = 3) 
  
# Legend
par(mar=c(0, 0, 0, 0))
plot.new()
legend("center", c(paste("All; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("All; ",levels(RDA$slope)[2]," Slope", sep=""),
      paste("Active; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("Active; ",levels(RDA$slope)[2]," Slope", sep="")),
      pt.lwd=2, col="black", pt.bg=c("wheat3", "green3", "wheat3",
      "green3"), pch=c(21,21,22,22), bty='n', ncol=4, cex=1.5, pt.cex=2)  

# Close Plot Device
dev.off()
graphics.off()
#```

## Show Plot
#```{r}
img <- readPNG("../figures/BIDE_Ordination.png")
grid.raster(img)
```



```{r}
ec.dbRDA.sim <- function(shared = " ", level = "0.03", design = " "){

  # Example Data
  #   shared     = "../microbide/SbyS/Condition1.txt"
  #   cutoff     = "0.03"
  #   design     = "../data/simmy.design.txt"

  source("../bin/DiversityFunctions.r")
  require("vegan")

  # Import Site by OTU Matrix
  ec_data.sim <- read.otu(shared, "0.03")
  design <- read.delim(design, header=T, row.names=1)

  # Remove OTUs with less than ten observations
  ec_data_red <- ec_data.sim[,colSums(ec_data.sim) >= 2]

  # Create factors for model
  slope <- design$slope       # factor 1
  molecule <- design$molecule # factor 2
  paired <- design$paired     # pairs
  site <- design$site         # sites
  station <- design$station

  # Create Y matrix with factor 1 and 2
  slope.molecule <- data.frame(cbind(as.character(slope),
    as.character(molecule)))

  # create unique treat ID vector
  slope.molecule.concat <- do.call(paste, c(slope.molecule[c("X1", "X2")],
    sep = ""))
  pair.station <- c(rep(seq(1:9), each=2), rep(seq(1:5), each=2),
    rep(seq(1:10), each=2), rep(seq(1:9), each=2))

  # Create a vector of molecules by station
  station.molecule.concat <- paste(station, molecule, sep = "")

  # Calculate Presense Absence
  dataPA <- (ec_data_red > 0)*1
  
  # Calculating Relative Abundance
  dataREL <- ec_data_red
  for(i in 1:nrow(ec_data_red)){
    dataREL[i,] = ec_data_red[i,]/sum(ec_data_red[i,])
  }
  
  # Log Transform Relative Abundance
  dataREL.l <- decostand(dataREL,method="log")

  # Chord Transformation
  dataREL.c <- decostand(dataREL, method="normalize")

  # Hellinger Transformation
  dataREL.h <- decostand(dataREL, method="hellinger")

  # Distance Based Redundancy Analysis
  dbRDA <- capscale(dataREL ~ slope + molecule , distance="bray")

#   head(summary(dbRDA))
#   anova(dbRDA, by="terms", permu=999)
#   varpart(dataREL,  ~ slope, ~ molecule)

  # Percent Variance Explained Using PCoA (Axis 1,2,3)
  explainvar1 <- round(dbRDA$CCA$eig[1]/sum(dbRDA$CCA$eig, dbRDA$CA$eig)*100,2)
  explainvar2 <- round(dbRDA$CCA$eig[2]/sum(dbRDA$CCA$eig, dbRDA$CA$eig)*100,2)

  #RDA <- as.data.frame(dbRDA$CCA$wa)
  RDA <- as.data.frame(scores(dbRDA, display = c("wa"), scaling = "sites"))
  RDA$molecule <- design$molecule
  RDA$slope <- design$slope
  RDA$labs <- slope.molecule.concat

  RDA.out <- list("RDA" = RDA,
                  "explainvar1" = explainvar1,
                  "explainvar2" = explainvar2)

  return(RDA.out)

}

```


