---
title: "Evolution Canyon"
author: "Mario E. Muscarella, Kenneth J. Locey, & Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output:
  pdf_document:
  fig_caption: true
---

Dormancy, dispersal, and environmental filtering at "Evolution Canyon"

Brief Description


# Initial Setup
```{r results='hide', message=FALSE}
rm(list=ls())
getwd()
setwd("~/GitHub/evolution-canyon/analyses")

# Import Required Packages
require("png")
require("grid")
require("vegan")

# Import Scripts
source("../bin/ec.dbRDA.R")

# Common Functions
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```


# Figure 1: Conceptual model/theory and predictions of how environmental filtering and dormancy influence community assembly (with dispersal being implicit and more or less high) 

**Jay is updating this**

```{r fig.width=2, fig.height=4,echo=FALSE,fig.cap="Figure 1: Conceptual Model"}
img <- readPNG("../figures/ConceptualModel.png")
grid.raster(img)
```


# Figure 2: Photo of EC and simulation of EC-like landscape

```{r fig.width=2, fig.height=4,echo=FALSE,fig.cap="Figure 1: Conceptual Model"}
img1 <- readPNG("../figures/ECLandscape.png")
img2 <- readPNG("../figures/BIDELandscape.png")
grid.newpage()
grid.raster(img1, x=0.5, y = 0.75, width=0.5)
grid.raster(img2, x = 0.5, y = 0.35, width=0.5)
```

# Figure 3: BIDE Model Ordinations

Ordination of simulated data capturing processes/scenarios described in Fig. 1. Here, we only have models 1, 2, 3, and 6; missing models 4 and 5. Can these other scenarios be captures with simplex?


```{r}
dir <- "../microbide/SbyS/"
input <- list.files(dir, "*.txt")
input <- input[c(1, 2, 3, 6)]

input
test <- read.delim(paste(dir, input[1], sep = ""), header = T)
rowSums(as.matrix(test[ 1, -c(1:3)]))

# Create Blank Output
dbRDA.output <- list()

# Run PCoA On All Samples
for (i in 1:length(input)){
  dbRDA.output[[input[i]]] = ec.dbRDA.sim(shared = paste(dir, input[i], sep=""), 
                                             level = "0.03", 
                                             design = "../data/simmy.design.txt")
}

png(filename = "../figures/BIDE_Ordination.png", 
    width = 1600, height = 1200, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0,0), oma=c(6, 5, 1, 1))
layout(matrix(c(5, 5, 1:4), 3, 2, byrow=T), heights=c(1, rep(5, 2)))

labs <- c("A", "B", "C", "D", "E", "F")
x.lab <- c(F, F, T, T)
y.lab <- c(T, F, T, F)

# Plot Each
for (p in 1:length(dbRDA.output)){
  RDA <- dbRDA.output[[p]]$RDA
  explainvar1 <- dbRDA.output[[p]]$explainvar1
  explainvar2 <- dbRDA.output[[p]]$explainvar2
  print(explainvar1)
  print(explainvar2)

  x.dim <- c(-0.6, 0.6)
  y.dim <- c(-0.5, 0.5)

  # Initiate Plot
  plot(RDA$CAP1, RDA$CAP2, xlab = "", ylab="",
    xlim=x.dim,ylim= y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)

  axis(side=1, las=1, label=x.lab[p], lwd.ticks = 2, cex.axis = 1.25)
  axis(side=2, las=1, label=y.lab[p], lwd.ticks = 2, cex.axis = 1.25)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  abline(h=0, lty="dotted")
  abline(v=0, lty="dotted")
  mol.shape <- rep(NA, dim(RDA)[1])
    for (i in 1:length(mol.shape)){
      if (RDA$molecule[i] == "DNA"){mol.shape[i] = 21}
      else {mol.shape[i] = 22}
      }
  slope.color <- rep(NA, dim(RDA)[1])
    for (i in 1:length(slope.color)){
      if (RDA$slope[i] == levels(RDA$slope)[1]) {slope.color[i] = "wheat3"}
      else {slope.color[i] = "green3"}
      }
  points(RDA$CAP1, RDA$CAP2, pch=mol.shape, cex=2.0, col="black", bg=slope.color, lwd=2)
  box(lwd=2)
  
  text(x = -0.58, y = 0.45, labs[p], cex = 1.5)
}
  
mtext("Axis 1", side = 1, cex = 1.5, outer=T, line = 3)
mtext("Axis 2", side = 2, cex = 1.5, outer=T, line = 3) 
  
# Legend
par(mar=c(0, 0, 0, 0))
plot.new()
legend("center", c(paste("All; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("All; ",levels(RDA$slope)[2]," Slope", sep=""),
      paste("Active; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("Active; ",levels(RDA$slope)[2]," Slope", sep="")),
      pt.lwd=2, col="black", pt.bg=c("wheat3", "green3", "wheat3",
      "green3"), pch=c(21,21,22,22), bty='n', ncol=4, cex=1.5, pt.cex=2)  

# Close Plot Device
dev.off()
graphics.off()
```

## Show Plot
```{r}
img <- readPNG("../figures/BIDE_Ordination.png")
grid.raster(img)
```

# Multivariate Analysis of BIDE Simulations

A general discussion that the simulations agreed with theory outlined in Fig. 1. 
Will reference ordinations and PCoA. 
Discussed Table 1 having summary data from the PERMANOVA results.

```{r}
input <- list.files("../microbide/SbyS/", "*.txt")
input <- input[c(1,2,3,6)]

permanova.output <- list()

for (i in c(1:length(input))){
  shared <- paste("../microbide/SbyS/", input[i], sep="")
  design <- "../data/simmy.design.txt"
  sbys.in <- read.otu(shared, "0.03")
  design.in <- read.delim(design, header=T, row.names=1)
  
  # Remove any OTUs with <2 OTUs
  sbys.red <- sbys.in[,which(colSums(sbys.in) >= 10)]
  
  # Relative Abundances
  sbysREL <- sbys.red
  for(j in 1:nrow(sbys.red)){
    sbysREL[j,] = sbys.red[j,]/sum(sbys.red[j,])
  }
  
  
  # Log Transformation
  sbysREL.log <- decostand(sbysREL, method="log")
  Adonis.b <- adonis(sbysREL.log ~ design.in$molecule * design.in$slope)
  permanova.output[[input[i]]] = Adonis.b
  }


```

# Figure 4: Emperical Data

Ordination of empirical data consistent with the view that non-random entrance and exiting from dormancy on top of environmental filtering influence community assembly

```{r}
shared <- "../data/EC.bac.final.shared"
design.in <- "../data/design.txt"
level  <-  "0.03"

# Run PCoA On Emperical Data

ec_data <- read.otu(shared, "0.03")

# Note: owing to amplification issues, we only sequenced 76 of 80 samples.
# The four samples not included are C-1E-R, EC-2G-R, EC-2J-R, EC-6I-D
dim(ec_data)
counts <- count.groups(otu.matrix = ec_data)
low.coverage <- counts[which(counts < 50000)]

# Remove low coverage samples 
ec_data2 <- ec_data[-c(which(rownames(ec_data) %in% c(names(low.coverage)))), ]

# Remove problematic samples (EC_2A_D, EC_2A_R, EC_2C_R, EC_2D_R)
#ec_data.tmp <- ec_data2[-c(9,20:21,24:27,32,37,74),]
ec_data3 <- ec_data2[-c(which(rownames(ec_data2) %in% 
                     c("EC_2A_D", "EC_2A_R","EC_2C_R", "EC_2D_R"))), ]

rownames(ec_data3)

# Remove OTUs with less than two observations across samples
ec_data_red <- ec_data3[,which(colSums(ec_data3) >= 2)]

# Import Design
design <- read.delim(design.in, header=T, row.names=1)

# Match OTU matrix and design
design_red <- design[which(rownames(design) %in% rownames(ec_data_red)),]
```

# Run dbRDA By Hand
```{r}
# Create factors for model
slope <- design_red$slope # factor 1
molecule <- design_red$molecule # factor 2
paired <- design_red$paired_across_slope
site <- design_red$site
station <- design_red$station
slope.molecule <- data.frame(cbind(as.character(slope),
                                   as.character(molecule))) # Y matrix with factor 1 and 2
slope.molecule.concat <- do.call(paste, c(slope.molecule[c("X1", "X2")],
                                          sep = "")) # create unique treat ID vector
# Create a vector of molecules by station
station.molecule.concat <- paste(station, molecule, sep = "")


pair.station <- c(rep(seq(1:9), each=2), rep(seq(1:5), each=2), 
                  rep(seq(1:10), each=2), rep(seq(1:9), each=2))
# To include "questionable" data uncomment the following
# pair.station <- c(rep(seq(1:9), each=2), rep(seq(1:8), each=2),
#    rep(seq(1:10), each=10), rep(seq(1:9), each=2))

# Define OTU Matrix
otu.matrix <- ec_data_red

# Calculate Presense Absence
dataPA <- (otu.matrix > 0)*1

# Calculating Relative Abundance
dataREL <- otu.matrix
for(i in 1:nrow(otu.matrix)){
  dataREL[i,] = otu.matrix[i,]/sum(otu.matrix[i,])
  }

# Log Transform Relative Abundance
dataREL.l <- decostand(dataREL,method="log")

# Chord Transformation
dataChord <- decostand(ec_data_red, method="normalize")

# Hellinger Transformation
dataHell <- decostand(ec_data_red, method="hellinger")

# Create Distance Matrix with bray (deafault), manhattan, euclidean, canberra, bray, kulczynski, jaccard, gower, altGower, morisita, horn, mountford, raup, binomial, or chao. Most should be part of vegan, but possilbly 'labdsv' or 'BiodiversityR' packages
samplePA.dist <- vegdist(dataREL.l,method="bray")
sampleREL.dist <- vegdist(dataREL.l,method="bray")

# Distance Based Redundancy Analysis
#dbRDA <- capscale(dataRel.l ~ slope + molecule + Condition(paired), distance="bray")
dbRDA <- capscale(dataREL.l ~ slope + molecule, distance="bray")


#   head(summary(dbRDA))
#   anova(dbRDA, by="terms", permu=999)
#   varpart(dataREL,  ~ slope, ~ molecule)

# Percent Variance Explained Using PCoA (Axis 1,2,3)
explainvar1 <- round(dbRDA$CCA$eig[1]/sum(dbRDA$CCA$eig, dbRDA$CA$eig)*100,2)
explainvar2 <- round(dbRDA$CCA$eig[2]/sum(dbRDA$CCA$eig, dbRDA$CA$eig)*100,2)

#RDA <- as.data.frame(dbRDA$CCA$wa)
RDA <- as.data.frame(scores(dbRDA, display = c("wa"), scaling = "sites"))
RDA$molecule <- molecule
RDA$slope <- slope
RDA$labs <- slope.molecule.concat

RDA.out <- list("RDA" = RDA,
                "explainvar1" = explainvar1,
                "explainvar2" = explainvar2)
```


```{r}
dbRDA.output = ec.dbRDA(otu.matrix = ec_data_red, design = design)


# Plot
png(filename = "../figures/EC_Ordination.png", 
    width = 1000, height = 1000, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0,0), oma=c(5, 5, 1, 1))
layout(matrix(c(2, 1), 2, 1, byrow=T), heights=c(0.5, 4.5))

x.lab <- c(T)
y.lab <- c(T)

# Plot
  RDA <- dbRDA.output$RDA
  explainvar1 <- dbRDA.output$explainvar1
  explainvar2 <- dbRDA.output$explainvar2

  x.dim <- c(-0.5, 0.5)
  y.dim <- c(-0.4, 0.3)

  # Initiate Plot
  plot(RDA$CAP1, RDA$CAP2, xlab = "", ylab="",
    xlim=x.dim,ylim= y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)

  axis(side=1, las=1, label=T, lwd.ticks = 2)
  axis(side=2, las=1, label=T, lwd.ticks = 2)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  abline(h=0, lty="dotted")
  abline(v=0, lty="dotted")
  mol.shape <- rep(NA, dim(RDA)[1])
    for (i in 1:length(mol.shape)){
      if (RDA$molecule[i] == "DNA"){mol.shape[i] = 21}
      else {mol.shape[i] = 22}
      }
  slope.color <- rep(NA, dim(RDA)[1])
    for (i in 1:length(slope.color)){
      if (RDA$slope[i] == levels(RDA$slope)[1]) {slope.color[i] = "wheat3"}
      else {slope.color[i] = "green3"}
      }
  points(RDA$CAP1, RDA$CAP2, pch=mol.shape, cex=1.5, col="black", bg=slope.color, lwd=2)
  ordiellipse(cbind(RDA$CAP1, RDA$CAP2), RDA$labs, kind="sd", conf=0.95,
  lwd=2, lty=3, draw = "lines", col = "black", label=FALSE)
  box(lwd=2)
  
mtext("Axis 1", side = 1, cex = 1.5, outer=T, line = 2.5)
mtext("Axis 2", side = 2, cex = 1.5, outer=T, line = 2.5) 
  
# Legend
par(mar=c(0, 0, 0, 0))
plot.new()
legend("center", c(paste("All; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("All; ",levels(RDA$slope)[2]," Slope", sep=""),
     paste("Active; ",levels(RDA$slope)[1]," Slope", sep=""),
   paste("Active; ",levels(RDA$slope)[2]," Slope", sep="")),
   pt.lwd=2, col="black", pt.bg=c("wheat3", "green3", "wheat3",
   "green3"), pch=c(21,21,22,22), bty='n', ncol=2, cex=0.8, pt.cex=1)  

# Close Plot Device
dev.off()
graphics.off()
#```

## Show Plot
#```{r}
img <- readPNG("../figures/EC_Ordination.png")
grid.raster(img)
```

## PCoA Loadings
```{r}



```

## PERMANOVA
```{r}
# Match OTU matrix and design
design_red <- design[which(rownames(design) %in% rownames(ec_data_red)),]


# Relative Abundances
sbysREL <- ec_data_red
for(j in 1:nrow(ec_data_red)){
  sbysREL[j,] = ec_data_red[j,]/sum(ec_data_red[j,])
}
  
# Log Transformation
sbysREL.log <- decostand(sbysREL, method="log")
Adonis.b <- adonis(sbysREL.log ~ design_red$molecule * design_red$slope)
Adonis.b
```


## Within Site Diversity
```{r}
EC.OTU <- read.otu(shared, "0.03")
EC.design <- read.delim(design, header=T, row.names=1)

# Remove Bad Samples
#EC.bad <- rownames(EC.OTU)[c(9,20:21,24:27,32,37,74)]

# Samples by molecules
DNA <- rownames(EC.design[which(EC.design$molecule == "DNA"), ])
RNA <- rownames(EC.design[which(EC.design$molecule == "RNA"), ])

par(mfrow = c(2, 1), mar = c(2, 2, 2, 1), oma = c(4, 3, 1, 1))
hist(rowSums(EC.OTU[which(rownames(EC.OTU) %in% DNA), ]),  las = 1, ylim = c(0, 18),
     main = "DNA", xlab = "", ylab = "", breaks = seq(0, 250000, by = 25000))
mtext(side = 2, "Sites", line = 3, cex = 1.5)
hist((rowSums(EC.OTU[which(rownames(EC.OTU) %in% RNA), ])),  las = 1, ylim = c(0, 18),
     main = "RNA",xlab = "", ylab = "", breaks = seq(0, 250000, by = 25000))
mtext(side = 1, "Counts", line = 3, cex = 1.5)
mtext(side = 2, "Sites", line = 3, cex = 1.5)

hist(log10(rowSums(EC.OTU[which(rownames(EC.OTU) %in% EC.bad), ])), main = "Bad")

hist(log10(rowSums(EC.OTU[-(which(rownames(EC.OTU) %in% EC.bad)), ])))

# Remove any OTUs with <2 occurences (within and across sites)
EC.OTU.2 <- EC.OTU[,which(colSums(EC.OTU) >= 2)]
dim(EC.OTU.2)

# Separate into DNA and RNA
EC.DNA <- EC.OTU.2[which(EC.design$molecule == "DNA"), ]
EC.RNA <- EC.OTU.2[which(EC.design$molecule == "RNA"), ]

# Function for Calculating Species Richness w/ Resampling
richness.iter <- function(otu.matrix  = " ", 
                          sample.size = " ", 
                          iters       = " "){
  rich.matrix <- matrix(NA, dim(otu.matrix)[1], iters)
  rownames(rich.matrix) <- rownames(otu.matrix)
  for (i in 1:iters){
    temp.matrix <- rrarefy(otu.matrix, sample.size)
    rich.matrix[,i] <- rowSums((temp.matrix>0)*1)
  }
  rich.mean <- apply(rich.matrix, 1, mean)
  rich.se <- apply(rich.matrix, 1, sem)
  return(data.frame(S = rich.mean, S_se = rich.se))
}

# Function for Simpsons Evenness w/ Resampling
evenness.iter <- function(otu.matrix  = " ", 
                          sample.size = " ", 
                          iters       = " "){
  simp_even <- function(SAD = " "){
  SAD <- subset(SAD, SAD > 0)
  S <- length(SAD); N <- sum(SAD); X <- rep(NA, S)
  for (i in 1:S){
    X[i] <- (SAD[i]*(SAD[i] - 1)) / (N * (N - 1))
    }
  D <- sum(X); e_d <- (1/D)/S
  return(e_d)
  }
  even.matrix <- matrix(NA, dim(otu.matrix)[1], iters)
  rownames(even.matrix) <- rownames(otu.matrix)
  for (i in 1:iters){
    temp.matrix <- rrarefy(otu.matrix, sample.size)
    even.matrix[,i] <- apply(temp.matrix, 1, simp_even)
    }
  even.mean <- apply(even.matrix, 1, mean)
  even.se <- apply(even.matrix, 1, sem)
  return(data.frame(E = even.mean, E_se = even.se))
}

S.iter <- richness.iter(EC.OTU.2, min(rowSums(EC.OTU.2) * 0.9), 100)
E.iter <- evenness.iter(EC.OTU.2, min(rowSums(EC.OTU.2) * 0.9), 100)

```





