---
title: "Evolution Canyon"
author: "Jay T. Lennon, Kenneth J. Locey, & Mario E. Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
   - \usepackage{graphics}
output:
  pdf_document:
  fig_caption: true
---

Dormancy, dispersal, and environmental filtering at "Evolution Canyon"

Brief Description


# Initial Setup
```{r results='hide', message=FALSE}
rm(list=ls())
getwd()
setwd("~/GitHub/evolution-canyon/analyses")

# Import Required Packages
require("png")
require("grid")
require("vegan")

# Import Scripts
source("../bin/ec.dbRDA.R")

# Common Functions
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```


# Figure 1: Conceptual model/theory and predictions of how environmental filtering and dormancy influence community assembly (with dispersal being implicit and more or less high) 

**Jay is updating this**

```{r fig.width=2, fig.height=4,echo=FALSE,fig.cap="Figure 1: Conceptual Model"}
img <- readPNG("../figures/ConceptualModel.png")
grid.raster(img)
```


# Figure 2: Photo of EC and simulation of EC-like landscape

```{r fig.width=2, fig.height=4,echo=FALSE,fig.cap="Figure 1: Conceptual Model"}
img1 <- readPNG("../figures/ECLandscape.png")
img2 <- readPNG("../figures/BIDELandscape.png")
grid.newpage()
grid.raster(img1, x=0.5, y = 0.75, width=0.5)
grid.raster(img2, x = 0.5, y = 0.35, width=0.5)
```

# Figure 3: BIDE Model Ordinations

Ordination of simulated data capturing processes/scenarios described in Fig. 1. Here, we only have models 1, 2, 3, and 6; missing models 4 and 5. Can these other scenarios be captures with simplex?


```{r}
dir <- "../microbide/SbyS/"
input <- list.files(dir, "*.txt")
input <- input[c(1, 2, 3, 6)]

# Create Blank Output
dbRDA.output <- list()

# Run PCoA On All Samples
for (i in 1:length(input)){
  dbRDA.output[[input[i]]] = ec.dbRDA.sim(shared = paste(dir, input[i], sep=""), 
                                             level = "0.03", 
                                             design = "../data/simmy.design.txt")
}

png(filename = "../figures/BIDE_Ordination.png", 
    width = 1600, height = 1200, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0,0), oma=c(6, 5, 1, 1))
layout(matrix(c(5, 5, 1:4), 3, 2, byrow=T), heights=c(1, rep(5, 2)))

labs <- c("A", "B", "C", "D", "E", "F")
x.lab <- c(F, F, T, T)
y.lab <- c(T, F, T, F)

# Plot Each
for (p in 1:length(dbRDA.output)){
  RDA <- dbRDA.output[[p]]$RDA
  explainvar1 <- dbRDA.output[[p]]$explainvar1
  explainvar2 <- dbRDA.output[[p]]$explainvar2

  x.dim <- c(-0.6, 0.6)
  y.dim <- c(-0.5, 0.5)

  # Initiate Plot
  plot(RDA$CAP1, RDA$CAP2, xlab = "", ylab="",
    xlim=x.dim,ylim= y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)

  axis(side=1, las=1, label=x.lab[p], lwd.ticks = 2, cex.axis = 1.25)
  axis(side=2, las=1, label=y.lab[p], lwd.ticks = 2, cex.axis = 1.25)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  abline(h=0, lty="dotted")
  abline(v=0, lty="dotted")
  mol.shape <- rep(NA, dim(RDA)[1])
    for (i in 1:length(mol.shape)){
      if (RDA$molecule[i] == "DNA"){mol.shape[i] = 21}
      else {mol.shape[i] = 22}
      }
  slope.color <- rep(NA, dim(RDA)[1])
    for (i in 1:length(slope.color)){
      if (RDA$slope[i] == levels(RDA$slope)[1]) {slope.color[i] = "wheat3"}
      else {slope.color[i] = "green3"}
      }
  points(RDA$CAP1, RDA$CAP2, pch=mol.shape, cex=2.0, col="black", bg=slope.color, lwd=2)
  box(lwd=2)
  
  text(x = -0.58, y = 0.45, labs[p], cex = 1.5)
}
  
mtext("Axis 2", side = 1, cex = 1.5, outer=T, line = 3)
mtext("Axis 1", side = 2, cex = 1.5, outer=T, line = 3) 
  
# Legend
par(mar=c(0, 0, 0, 0))
plot.new()
legend("center", c(paste("All; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("All; ",levels(RDA$slope)[2]," Slope", sep=""),
      paste("Active; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("Active; ",levels(RDA$slope)[2]," Slope", sep="")),
      pt.lwd=2, col="black", pt.bg=c("wheat3", "green3", "wheat3",
      "green3"), pch=c(21,21,22,22), bty='n', ncol=4, cex=1.5, pt.cex=2)  

# Close Plot Device
dev.off()
graphics.off()
```

## Show Plot
```{r}
img <- readPNG("../figures/BIDE_Ordination.png")
grid.raster(img)
```

# Multivariate Analysis of BIDE Simulations

A general discussion that the simulations agreed with theory outlined in Fig. 1. 
Will reference ordinations and PCoA. 
Discussed Table 1 having summary data from the PERMANOVA results.

```{r}
input <- list.files("../microbide/SbyS/", "*.txt")
input <- input[c(1,2,3,6)]

permanova.output <- list()

for (i in c(1:length(input))){
  shared <- paste("../microbide/SbyS/", input[i], sep="")
  design <- "../data/simmy.design.txt"
  sbys.in <- read.otu(shared, "0.03")
  design.in <- read.delim(design, header=T, row.names=1)
  
  # Remove any OTUs with <2 OTUs
  sbys.red <- sbys.in[,which(colSums(sbys.in) >= 10)]
  
  # Relative Abundances
  sbysREL <- sbys.red
  for(j in 1:nrow(sbys.red)){
    sbysREL[j,] = sbys.red[j,]/sum(sbys.red[j,])
  }
  
  
  # Log Transformation
  sbysREL.log <- decostand(sbysREL, method="log")
  Adonis.b <- adonis(sbysREL.log ~ design.in$molecule * design.in$slope)
  permanova.output[[input[i]]] = Adonis.b
  }


```

# Figure 4: Emperical Data

Ordination of empirical data consistent with the view that non-random entrance and exiting from dormancy on top of environmental filtering influence community assembly

```{r}
shared <- "../data/EC.bac.final.shared"
design <- "../data/design.txt"
level  <-  "0.03"

# Run PCoA On Emperical Data
dbRDA.output = ec.dbRDA(shared = shared, level = level, design = design)

png(filename = "../figures/EC_Ordination.png", 
    width = 1000, height = 1000, res = 96 * 2)

# Plot Parameters
par(mfrow=c(1,1), mar=c(0.5,0.5,0,0), oma=c(5, 5, 1, 1))
layout(matrix(c(2, 1), 2, 1, byrow=T), heights=c(0.5, 4.5))

x.lab <- c(T)
y.lab <- c(T)

# Plot
  RDA <- dbRDA.output$RDA
  explainvar1 <- dbRDA.output$explainvar1
  explainvar2 <- dbRDA.output$explainvar2

  x.dim <- c(-0.6, 0.6)
  y.dim <- c(-0.5, 0.5)

  # Initiate Plot
  plot(RDA$CAP1, RDA$CAP2, xlab = "", ylab="",
    xlim=x.dim,ylim= y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)

  axis(side=1, las=1, label=T, lwd.ticks = 2)
  axis(side=2, las=1, label=T, lwd.ticks = 2)
  axis(side = 1, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1) 
  axis(side = 2, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  axis(side = 3, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)     
  axis(side = 4, lwd.ticks = 2, tck=0.01, labels = F,  cex.axis = 2, las = 1)
  abline(h=0, lty="dotted")
  abline(v=0, lty="dotted")
  mol.shape <- rep(NA, dim(RDA)[1])
    for (i in 1:length(mol.shape)){
      if (RDA$molecule[i] == "DNA"){mol.shape[i] = 21}
      else {mol.shape[i] = 22}
      }
  slope.color <- rep(NA, dim(RDA)[1])
    for (i in 1:length(slope.color)){
      if (RDA$slope[i] == levels(RDA$slope)[1]) {slope.color[i] = "wheat3"}
      else {slope.color[i] = "green3"}
      }
  points(RDA$CAP1, RDA$CAP2, pch=mol.shape, cex=1.5, col="black", bg=slope.color, lwd=2)
  box(lwd=2)
  
mtext("Axis 2", side = 1, cex = 1.5, outer=T, line = 2.5)
mtext("Axis 1", side = 2, cex = 1.5, outer=T, line = 2.5) 
  
# Legend
par(mar=c(0, 0, 0, 0))
plot.new()
legend("center", c(paste("All; ",levels(RDA$slope)[1]," Slope", sep=""),
      paste("All; ",levels(RDA$slope)[2]," Slope", sep=""),
     paste("Active; ",levels(RDA$slope)[1]," Slope", sep=""),
   paste("Active; ",levels(RDA$slope)[2]," Slope", sep="")),
   pt.lwd=2, col="black", pt.bg=c("wheat3", "green3", "wheat3",
   "green3"), pch=c(21,21,22,22), bty='n', ncol=4, cex=0.6, pt.cex=0.8)  

# Close Plot Device
dev.off()
graphics.off()
```

## Show Plot
```{r}
img <- readPNG("../figures/EC_Ordination.png")
grid.raster(img)
```

## PCoA Loadings
```{r}



```

## PERMANOVA
```{r}
shared <- "../data/EC.bac.final.shared"
design <- "../data/design.txt"

sbys.in <- read.otu(shared, "0.03")
design.in <- read.delim(design, header=T, row.names=1)
  
# Remove any OTUs with <2 OTUs
sbys.red <- sbys.in[,which(colSums(sbys.in) >= 10)]
  
# Relative Abundances
sbysREL <- sbys.red
for(j in 1:nrow(sbys.red)){
  sbysREL[j,] = sbys.red[j,]/sum(sbys.red[j,])
}
  
# Log Transformation
sbysREL.log <- decostand(sbysREL, method="log")
Adonis.b <- adonis(sbysREL.log ~ design.in$molecule * design.in$slope)

```


## Within Site Diversity
```{r}
EC.OTU <- read.otu(shared, "0.03")
EC.design <- read.delim(design, header=T, row.names=1)

# Remove Bad Samples
EC.bad <- rownames(EC.OTU)[c(9,20:21,24:27,32,37,74)]

# Samples by molecules
DNA <- rownames(EC.design[which(EC.design$molecule == "DNA"), ])
RNA <- rownames(EC.design[which(EC.design$molecule == "RNA"), ])

hist(log10(rowSums(EC.OTU[which(rownames(EC.OTU) %in% DNA), ])), main = "DNA")
hist(log10(rowSums(EC.OTU[which(rownames(EC.OTU) %in% RNA), ])), main = "RNA")
hist(log10(rowSums(EC.OTU[which(rownames(EC.OTU) %in% EC.bad), ])), main = "Bad")

hist(log10(rowSums(EC.OTU[-(which(rownames(EC.OTU) %in% EC.bad)), ])))

# Remove any OTUs with <2 occurences (within and across sites)
EC.OTU.2 <- EC.OTU[,which(colSums(EC.OTU) >= 2)]
dim(EC.OTU.2)

# Separate into DNA and RNA
EC.DNA <- EC.OTU.2[which(EC.design$molecule == "DNA"), ]
EC.RNA <- EC.OTU.2[which(EC.design$molecule == "RNA"), ]

# Function for Calculating Species Richness w/ Resampling
richness.iter <- function(otu.matrix  = " ", 
                          sample.size = " ", 
                          iters       = " "){
  rich.matrix <- matrix(NA, dim(otu.matrix)[1], iters)
  rownames(rich.matrix) <- rownames(otu.matrix)
  for (i in 1:iters){
    temp.matrix <- rrarefy(otu.matrix, sample.size)
    rich.matrix[,i] <- rowSums((temp.matrix>0)*1)
  }
  rich.mean <- apply(rich.matrix, 1, mean)
  rich.se <- apply(rich.matrix, 1, sem)
  return(data.frame(S = rich.mean, S_se = rich.se))
}

# Function for Simpsons Evenness w/ Resampling
evenness.iter <- function(otu.matrix  = " ", 
                          sample.size = " ", 
                          iters       = " "){
  simp_even <- function(SAD = " "){
  SAD <- subset(SAD, SAD > 0)
  S <- length(SAD); N <- sum(SAD); X <- rep(NA, S)
  for (i in 1:S){
    X[i] <- (SAD[i]*(SAD[i] - 1)) / (N * (N - 1))
    }
  D <- sum(X); e_d <- (1/D)/S
  return(e_d)
  }
  even.matrix <- matrix(NA, dim(otu.matrix)[1], iters)
  rownames(even.matrix) <- rownames(otu.matrix)
  for (i in 1:iters){
    temp.matrix <- rrarefy(otu.matrix, sample.size)
    even.matrix[,i] <- apply(temp.matrix, 1, simp_even)
    }
  even.mean <- apply(even.matrix, 1, mean)
  even.se <- apply(even.matrix, 1, sem)
  return(data.frame(E = even.mean, E_se = even.se))
}

S.iter <- richness.iter(EC.OTU.2, min(rowSums(EC.OTU.2) * 0.9), 100)
E.iter <- evenness.iter(EC.OTU.2, min(rowSums(EC.OTU.2) * 0.9), 100)

```





